---
title: "Exam 3"
author: "Emmaline Saunders"
date: "2024-11-26"
output: html_document
---

### Reading in the required libraries

```{r}
library(tidyverse)
library(janitor)
library(broom)
```

## Part 1

### Reading in the factuly salaries data

```{r}
fs_data <- read.csv("./FacultySalaries_1995.csv")
```

### Exploring the dataset

```{r}
head(fs_data)
glimpse(fs_data)
View(fs_data)
names(fs_data)
```
It's terribly messy. It definitely needs some scrub a dub dub action. 
 
### Cleaning the dataset

#### Pivoting the data longer
```{r}

fs_data_clean <- fs_data %>%
  pivot_longer(cols = c("AvgFullProfSalary", "AvgAssocProfSalary", 
                        "AvgAssistProfSalary", "AvgProfSalaryAll"),
               names_to = "rank_salary", 
               values_to = "salary") %>%
  pivot_longer(cols = c("AvgFullProfComp", "AvgAssocProfComp", 
                        "AvgAssistProfComp", "AvgProfCompAll"),
               names_to = "rank_comp", 
               values_to = "comp") %>%
  pivot_longer(cols = c("NumFullProfs", "NumAssocProfs", 
                        "NumAssistProfs", "NumInstructors", "NumFacultyAll"),
               names_to = "instructor_type", 
               values_to = "number_faculty")
```

#### Cleaing up data in the columns

```{r}
##rank and salary
fs_data_clean$rank_salary <- gsub("Prof", "", fs_data_clean$rank_salary)
fs_data_clean$rank_salary <- gsub("Avg", "", fs_data_clean$rank_salary)
fs_data_clean$rank_salary <- gsub("Salary", "", fs_data_clean$rank_salary)

##rank and comp
fs_data_clean$rank_comp <- gsub("Prof", "", fs_data_clean$rank_comp)
fs_data_clean$rank_comp <- gsub("Avg", "", fs_data_clean$rank_comp)
fs_data_clean$rank_comp <- gsub("Comp", "", fs_data_clean$rank_comp)

```
Presenting the clean data!
```{r}
print(fs_data_clean)
```


### Plotting the data
but first, I need to do something to the dataset...We still have some pesky values showing up
```{r}
filtered_fs_data <- fs_data_clean %>%
  filter(rank_salary != "All") %>%
  filter(Tier != "VIIB")
```

finally onto my favorite part...the plot!

```{r}
filtered_fs_data %>%
  ggplot(aes(x = rank_salary, y = salary, fill = rank_salary)) +
  geom_boxplot()+
  facet_wrap(~Tier) +
  theme_minimal() +
  labs(
    x = "Rank",           
    y = "Salary",
    fill = "Rank"
  )

```


## Part 2

### Creating the ANOVA Model

```{r}
anova_model <- aov(salary ~ State + Tier + rank_salary, data = filtered_fs_data)

```

but what does the model tell us?

```{r}
summary(anova_model)
```

## Part 3

### Reading in the Juniper Oils data

```{r}
jo_data <- read.csv("./Juniper_Oils.csv") 
```

### Exploring the data
```{r}

head(jo_data)
glimpse(jo_data)
View(jo_data)
names(jo_data)

```

Ugh such a messy dataset. Time for some elbow grease!

### Cleaning the dataset

```{r}
tidy_jo_data <- jo_data %>%
  pivot_longer(cols = c("alpha.pinene","para.cymene","alpha.terpineol",
                        "cedr.9.ene","alpha.cedrene","beta.cedrene",
                        "cis.thujopsene","alpha.himachalene","beta.chamigrene",
                        "cuparene","compound.1","alpha.chamigrene",
                        "widdrol","cedrol","beta.acorenol",
                        "alpha.acorenol","gamma.eudesmol","beta.eudesmol",
                        "alpha.eudesmol","cedr.8.en.13.ol","cedr.8.en.15.ol",
                        "compound.2","thujopsenal"), 
               names_to = "chem_type", 
               values_to = "ms_conc") %>%
  mutate(chem_type = gsub("\\.", "-", chem_type))

print(tidy_jo_data)
```

Now that's a pristine dataset! I think we need to plot it and help it say something

## Part 4

### Plotting the data
```{r}
ggplot(tidy_jo_data, aes(x = YearsSinceBurn, y = ms_conc)) +
  geom_smooth() +
  facet_wrap(~ chem_type, scales = "free_y") +
  labs(
    x = "Years Since Burn",
    y = "Concentration"
  ) +
  theme_minimal()
```
What a beautiful plot!

## Part 5

What is the utility of a plot if we don't have statistics to see which variables are actually significant?

### Filtering and selecting the important stuff
```{r}
sig_chem <- tidy_jo_data %>%
  group_by(chem_type) %>%                       
  do({
    model <- glm(ms_conc ~ YearsSinceBurn, data = ., family = gaussian())
    tidy(model)                                    
  }) %>%
  ungroup() %>%
  filter(term == "YearsSinceBurn", p.value < 0.05) 
```

Great! now let's actually see what affects it!

```{r}
print(sig_chem)
```
